It's easier to move sideways

### 2023-09-01 manuscript.Rmd
multiple updates, including caching code chunks

---
title: "It's easier to move sideways"
author:
  - Kaleonani K.C. Hurley:
      institute: 
        - himb
        - nerr
      equal_contributor: yes
      correspondence: yes
      email: "hurleyk@hawaii.edu"
  - Mykle L. Hoban:
      institute:
        - himb
      equal_contributor: yes
  - Molly A. Timmers:
      institute: [himb, natgeo]
  - Kerry Reardon:
      institute: noaa
  - Robert J. Toonen:
      institute: himb
institute:
  - himb: Hawai&#x02BB;i Institute of Marine Biology
  - nerr: He&#x02BB;eia National Estuarine Research Reserve
  - natgeo: National Geographic Pristine Seas
  - noaa: National Oceanic and Atmospheric Administration
abstract:
  Shallow coral reefs are extensively studied, and although scleractinian corals have been recorded to 165 m, little is known about other mesophotic coral reef ecosystem (MCE) inhabitants. Brachyuran crabs fill many ecological and trophic niches on reefs, making them ideal candidates for evaluating species composition among depths. Here we ask if MCEs host the same communities as the shallower reefs spread across the entire 2400 km Hawaiian Archipelago. We deployed Autonomous Reef Monitoring Structures (ARMS) for two years among shallow sites (12 m) across the latitudinal gradient of the Hawaiian Archipelago to compare directly with a depth gradient (12, 30, 60, and 90 m) south of Oʻahu island. A total of 2284 individual crabs representing 90 morphospecies (19 families) were found. Assemblage composition was not statistically significantly different among shallow sites spread across the archipelago but was significantly different and highly stratified across the depth gradient.  We show that 90 m of depth is a more potent determinant of brachyuran community composition than the latitudinal, geographic (high vs low islands), or anthropogenic impact (human population size) gradient across the archipelago. Deeper reefs host significantly different brachyuran communities than shallow ones in Hawai&#x02BB;i, and only 3 of 90 morphospecies (~3%) occurring across the entire depth range (mesophotic to shallow) were sampled. Further research on the cryptic reef-dwelling fauna is needed for identifying areas of unique biodiversity, as well as boundaries between depth-stratified reef ecosystems.
bibliography: references/citations.bib
csl: references/vancouver.csl
output:
  word_document:
    pandoc_args:
      - "--filter=resources/pandoc-crossref"
      - "--lua-filter=resources/scholarly-metadata.lua"
      - "--lua-filter=resources/author-info-blocks.lua"
      - "--reference-doc=resources/ref.docx"
    keep_md: true
tblPrefix: Table
figPrefix:
  - "Fig."
  - "Figs."
---

### Keywords
<div custom-style="mono">ARMs, mesophotic coral ecosystems, community structure, dispersal, Hawai&#x02BB;i, brachyuran crabs</div>
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path="output/figures/fig-",
  dev="svg",
  dpi=300,
  message=FALSE,
  warning=FALSE
)

library(pander)
library(fs)
library(here)
library(english)

source("crabs.R")
if (!exists("crabs_done")) {
  crabs_done <<- FALSE
}
if (!crabs_done) {
  cc <<- setup_crabs()
  crabs_done <<- TRUE
}

@@

top_list <- function(species,n,units,len=4) {
  s1 <- map_chr(species,format_species)
  s2 <- map2_chr(n,units,~{
    str_glue("(with {.x} individuals on {.y} units)")
  })
  ss <- str_c(s1," ",s2)
  ss[1:(len-1)] %>%
    str_c(collapse=", ") %>%
    str_c(", and ",ss[len])
}

print_table <- function(tbl,id="",caption="") {
  if (!grepl("^: ",caption)) {
    caption <- paste0(": ",caption)
  }

@@

@@

all_alpha <- alpha_diversity(cc$crabs_untransformed)
deep_alpha <- alpha_diversity(cc$crabs_deep)
shallow_alpha <- alpha_diversity(cc$crabs_shallow)
```

# Introduction

# Materials and Methods

## Sample Collection

Brachyuran crab assemblages were sampled from benthic coral reef habitat in the Hawaiian Archipelago using Autonomous Reef Monitoring Structures (ARMS). ARMS are standardized collection devices designed to mimic coral reef structural complexity to attract settlement of reef cryptofauna [@brainardAutonomousReefMonitoring2009; @zimmermanArtificialReefMatrix2004]. They are constructed from 10 gray type 1 PVC plates (23 cm $\times$ 23 cm) stacked in an alternating series of open and semi-enclosed layers (figure xxx, ARMS unit). The stack is fastened to a larger base plate (35 cm $\times$ 45 cm) which can be attached to the benthos upon deployment [@brainardAutonomousReefMonitoring2009; @hurleyAssessmentShallowMesophotic2016; @knowltonCoralReefBiodiversity2010; @zimmermanArtificialReefMatrix2004].

@@

```{r arms-table,results="asis"}
arms_schedule <- cc$crabs_untransformed %>%
  sample_tibble() %>%
  separate_wider_regex(sample,patterns=c(recovery_year="^[0-9]+","_",sample=".*"))  %>%
  select(sample,everything()) %>%
  unite("range",deployment_year,recovery_year,sep="--") %>%
  mutate(
    island = case_when(
      str_remove_all(hawaiian_name,'[^a-zA-Z ]') != island ~ as.character(str_glue("{hawaiian_name} ({island})")),
      TRUE ~ as.character(hawaiian_name)
    ),
    island = fct_reorder(island,-lat)
  ) %>%
  count(island,range) %>%
  pivot_wider(names_from="range",values_from="n")  %>%
  select(island,order(colnames(.))) %>%
  mutate(across(-island,~replace_na(.x,0)))
print_table(arms_schedule,"arms_table","ARMS unit sampling schedule (number of units)")  
# print_table(here("data","tables","arms_deployment.csv"),"arms_table","ARMS unit sampling schedule (number of units)")
```

```{r site-map, fig.cap='Sampling map (Papahānaumokuākea Marine National Monument shaded in light blue)', results="asis", dev="svg", fig.width=9, fig.height=6}
library(sf)
library(rnaturalearth)

@@

plotz <- ggplot() +
  # geom_contour(data=bath_proj,mapping=aes(x=V1,y=V2,z = V3),  color="#cecece") +
  # geom_raster(data=bpr,aes(x=x,y=y,fill=z)) +
  # geom_tile(data=bath_proj,mapping=aes(x=V1,y=V2,fill = V3)) +
  # scale_fill_gradient2(low="dodgerblue4", mid="gainsboro", high="darkgreen") +
  geom_sf(data=hawaii %>% st_transform(transformer), color="black", fill="green", linewidth=0.4) +
  scale_fill_brewer(palette="Accent",name="Island") +
  geom_sf(data=pmnm,fill="#6ed8f6",alpha=0.1) +
  geom_sf(data=crab_sites %>% st_transform(transformer),aes(fill=hawaiian_name),size=3,shape=21,color="black") +
  coord_sf(xlim=box_after[c(1,3)],ylim=box_after[c(2,4)]) +
  scale_x_continuous(breaks = -seq(180, 155, -5)) +
  theme_minimal() +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(legend.text = element_text(size=10),legend.title = element_text(size=12))
# labela <- tribble(
#   ~lon,~lat,~text,
#   -170,25.4,"Paphānaumokuākea Marine\nNational Monument\n(NWHI)"
# ) %>%
#   st_as_sf(coords=c("lon","lat"),crs=4326, remove=F)  %>%
#   st_transform(transformer)
# 
# plotz +
#   geom_sf_text(aes(label=text),data=labela,size=3,angle=-17.0)
plotz
cat("{#fig:site_map}")
```

ARMS units were retrieved by encapsulating them in-situ in a mesh-lined crate (100 μm pore size) to prevent escape of motile organisms. Once returned to ship or shore, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate. Seawater surrounding disassembled units was sieved through 2 mm, 500 μm, and 100 μm mesh-size geologic sieves [see @lerayDNABarcodingMetabarcoding2015 for processing details], and retained macroorganisms were preserved in 95% ethanol. All adult brachyuran crabs ($\geq$ 5 mm) were photographed, identified to morphospecies, and preserved in 95% ethanol (Fig. 2d). Individual preserved specimens were accessioned into the invertebrate zoology collections at the Florida Museum of Natural History (catalog numbers xxx–xxx ).

## Data analysis

@@

## Environmental variables

Potential environmental factors affecting the community structure of brachyuran crabs on shallow reefs were assembled from published literature [@selkoeDNACoralReef2016] and the Pacific Islands Ocean Observing System (PacIOOS) ([@tbl:env_table]). Environmental factors were assembled only for shallow sites sampled during RAMP cruises.

```{r env-table,results="asis"}
print_table(here("data","tables","environmental_variables.csv"),"env_table","Potential environmental drivers of crab community structure, assembled from PacIOOS and @selkoeDNACoralReef2016")
```

# Results

## Diversity summary

### All ARMS

The `r nsamples(cc$crabs_untransformed)` ARMS units deployed across the Hawaiian Islands (in `r format_range(cc$crab_data$depth)` m depth) yielded a total of `r num(sum(sample_sums(cc$crabs_untransformed)),1)` adult crabs. Sampling units contained `r num(all_summary$abundance_mean,0.1)` $\pm$ `r num(all_summary$abundance_sd,0.1)` (mean $\pm$ SD) individuals per unit, and abundance ranged from `r format_range(all_summary$abundance_range)` individuals. Crabs were identified to `r sum(unlist(all_summary$types_breakdown))` distinct types (including `r all_summary$types_breakdown$species` species and `r all_summary$types_breakdown$morphospecies` morphospecies) in `r all_summary$taxonomy_breakdown$genera` genera and `r all_summary$taxonomy_breakdown$families` families, which accounted for ~36% of the known crab diversity reported from Hawaiian waters [@castroCatalogAnomuranBrachyuran2011]. The four most common species were: `r with(all_summary,top_list(top5,top5_n,top5_units))`. `r str_to_sentence(as.english(all_summary$singleton_count))` species occurred in a single unit across all sites (singletons), accounting for ~`r num(all_summary$singleton_count/ntaxa(cc$crabs_untransformed)*100,1)`% of total recovered crab species.

### Mesophotic ARMS

@@

## Alpha diversity

### Summary

@@

### Comparisons

For the complete dataset, there was no significant difference in observed species richness between the MHI and NWHI `r t_test(all_alpha$t_r_island)`, but Simpson diversity was significantly different, although mean values were similar with `r num(all_alpha$richness_main$richness,0.1)` $\pm$ `r num(all_alpha$richness_main$richness_sd,0.1)` in the MHI and `r num(all_alpha$richness_nwhi$richness,0.1)` $\pm$ `r num(all_alpha$richness_nwhi$richness_sd,0.1)` in the NWHI ([@fig:alpha_box]; `r t_test(all_alpha$t_s_island,plain=TRUE)`). When only shallow arms \< 30 m were considered, neither comparison was significant (richness: `r t_test(shallow_alpha$t_r_island,plain=TRUE)`; Simpson: `r t_test(shallow_alpha$t_s_island,plain=TRUE)`)<!--(richness: t = 0.77, DF = 94.47, p = 0.44; Simpson: t  = 1.93, DF = 84.76, p = 0.06)-->.

```{r alpha-box, fig.cap='Hello Dolly', results="asis", dev="svg", fig.width=10, fig.height=8}
ggplot(all_alpha$richness_table) + 
  geom_boxplot(aes(x=island_group,y=Simpson,fill=island_group)) +
  scale_fill_brewer(palette="Set2") +
  theme_bw() + 
  theme(panel.grid = element_blank())
cat("{#fig:alpha_box}")
```

### Effects of environmental variables

# Discussion

# References

@@

>>>>

---
title: "It's easier to move sideways"
author:
  - Kaleonani K.C. Hurley:
      institute:
        - himb
        - nerr
      equal_contributor: yes
      correspondence: yes
      email: "hurleyk@hawaii.edu"
  - Mykle L. Hoban:
      institute:
        - himb
      equal_contributor: yes
  - Molly A. Timmers:
      institute:
        - himb
        - natgeo
  - Kerry Reardon:
      institute: noaa
  - Robert J. Toonen:
      institute: himb
institute:
  - himb: Hawai&#x02BB;i Institute of Marine Biology
  - nerr: He&#x02BB;eia National Estuarine Research Reserve
  - natgeo: National Geographic Pristine Seas
  - noaa: National Oceanic and Atmospheric Administration
abstract:
  Shallow coral reefs are extensively studied, and although scleractinian corals have been recorded to 165 m, little is known about other mesophotic coral reef ecosystem (MCE) inhabitants. Brachyuran crabs fill many ecological and trophic niches on reefs, making them ideal candidates for evaluating species composition among depths. Here we ask if MCEs host the same communities as the shallower reefs spread across the entire 2400 km Hawaiian Archipelago. We deployed Autonomous Reef Monitoring Structures (ARMS) for two years among shallow sites (12 m) across the latitudinal gradient of the Hawaiian Archipelago to compare directly with a depth gradient (12, 30, 60, and 90 m) south of Oʻahu island. A total of 2284 individual crabs representing 90 morphospecies (19 families) were found. Assemblage composition was not statistically significantly different among shallow sites spread across the archipelago but was significantly different and highly stratified across the depth gradient.  We show that 90 m of depth is a more potent determinant of brachyuran community composition than the latitudinal, geographic (high vs low islands), or anthropogenic impact (human population size) gradient across the archipelago. Deeper reefs host significantly different brachyuran communities than shallow ones in Hawai&#x02BB;i, and only 3 of 90 morphospecies (~3%) occurring across the entire depth range (mesophotic to shallow) were sampled. Further research on the cryptic reef-dwelling fauna is needed for identifying areas of unique biodiversity, as well as boundaries between depth-stratified reef ecosystems.
bibliography: references/citations.bib
csl: references/vancouver.csl
output:
  word_document:
    pandoc_args:
      - "--toc-depth=2"
      - "--filter=resources/pandoc-crossref"
      - "--lua-filter=resources/scholarly-metadata.lua"
      - "--lua-filter=resources/author-info-blocks.lua"
      - "--reference-doc=resources/ref.docx"
    keep_md: true
tblPrefix: Table
figPrefix:
  - "Fig."
  - "Figs."
chapters: True
chaptersDepth: 1
chapDelim: ""
---

### Keywords {-}
<div custom-style="mono">ARMs, mesophotic coral ecosystems, community structure, dispersal, Hawai&#x02BB;i, brachyuran crabs</div>

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path="output/figures/fig-",
  dev="svg",
  dpi=300,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  cache.path = "cache/",
  cache.comments=FALSE
)

library(tidyverse)
library(pander)
library(fs)
library(here)
library(english)
library(phyloseq) 
library(vegan)
library(ggrepel)
library(here)
library(worrms)
```

```{r init, include=FALSE}
source("crabs.R")
if (!exists("crabs_done")) {
  crabs_done <<- FALSE
}
if (!crabs_done) {
  cc <<- setup_crabs()
  crabs_done <<- TRUE
}

@@

pval <- function(p) {
  case_when(
    p < 0.001 ~ "*p* \\< 0.001",
    p < 0.01 ~ "*p* \\< 0.01",
    TRUE ~ as.character(str_glue("*p* = {round(p,2)}"))
  )
}

permanova <- function(modl,d=1) {
  m <- suppressWarnings(broom::tidy(modl)) %>%
    filter(!term %in% c("Total","Residual"))
  if (nrow(m) > 1) {
    m %>%
      pmap_chr(~{
        with(list(...), {
          str_glue("{term}: DF = {round(df,1)}, SS = {round(SumOfSqs,1)}, $\\mathrm{{R}}^2$ = {round(R2,1)}, pseudo-F = {round(statistic,1)}, {pval(p.value)}")
        })
      }) %>%
      str_c(collapse = "; ")
  } else {
    with(m %>% slice(1),{
      str_glue("DF = {round(df,1)}, SS = {round(SumOfSqs,1)}, $\\mathrm{{R}}^2$ = {round(R2,1)}, pseudo-F = {round(statistic,1)}, {pval(p.value)}")
    }) 
  }
}

top_list <- function(species,n,units,len=4,fig=NULL) {
  s1 <- map_chr(species,format_species)
  s2 <- map2_chr(n,units,~{
    if (is.null(fig)) {
      str_glue("(with {.x} individuals on {.y} units)")
    } else {
      str_glue("(with {.x} individuals on {.y} units; [@fig:{fig}[]{{{{}}}}{{.y}}])")
    }
  })
  if (!is.null(fig)) {
    s2 <- map2_chr(s2,LETTERS[seq(length(s2))],~str_glue(.x))
  }
  ss <- str_c(s1," ",s2)
  ss[1:(len-1)] %>%
    str_c(collapse=", ") %>%
    str_c(", and ",ss[len])
}

print_table <- function(tbl,id="",caption="",hdr_case="sentence") {
  if (!grepl("^: ",caption)) {
    caption <- paste0(": ",caption)
  }

@@

  # can also use rename_with(str_to_sentence)
  names(tbl) <- switch(
    str_to_lower(hdr_case),
    sentence = str_to_sentence(names(tbl)),
    title = str_to_title(names(tbl)),
    names(tbl)
  )

@@

all_alpha <- alpha_diversity(cc$crabs_untransformed)
deep_alpha <- alpha_diversity(cc$crabs_deep)
shallow_alpha <- alpha_diversity(cc$crabs_shallow)

# palette for island/group colors
island_pal <- c(
  "Hawaii Island" = "#8a3ffc",
  "Hawai‘i Island" = "#8a3ffc",
  "Maui" = "#007d79",
  "Oahu" = "#ff7eb6",
  "O‘ahu" = "#ff7eb6",
  "Kauai" = "#fa4d56",
  "Kaua‘i" = "#fa4d56",

  "French Frigate Shoals" = "#fff1f1",
  "Lalo" = "#fff1f1",
  "Lisianski" = "#6fdc8c",
  "Kapou" = "#6fdc8c",
  "Pearl and Hermes Reef" = "#4589ff",
  "Manawai" = "#4589ff",
  "Kure Atoll" = "#d12771",
  "Hōlanikū" = "#d12771"
)
island_group_pal <- c(
  "main" = "#ffb400",
  "Main" = "#ffb400",
  "Northwest" = "#9080ff",
  "northwest" = "#9080ff"
)

```

# Introduction {-}

# Materials and Methods {-}

## Sample Collection {-}

Brachyuran crab assemblages were sampled from benthic coral reef habitat in the Hawaiian Archipelago using Autonomous Reef Monitoring Structures (ARMS). ARMS are standardized collection devices designed to mimic coral reef structural complexity to attract settlement of reef cryptofauna [@brainardAutonomousReefMonitoring2009; @zimmermanArtificialReefMatrix2004]. They are constructed from 10 gray type 1 PVC plates (23 cm $\times$ 23 cm) stacked in an alternating series of open and semi-enclosed layers ([@fig:fig_arms]). The stack is fastened to a larger base plate (35 cm $\times$ 45 cm) which can be attached to the benthos upon deployment [@brainardAutonomousReefMonitoring2009; @hurleyAssessmentShallowMesophotic2016; @knowltonCoralReefBiodiversity2010; @zimmermanArtificialReefMatrix2004].

![An Autonomous reef monitoring structure (ARMS) unit deployed on a reef in the Red Sea (photo credit: Michael Berumen, used with permission)](data/images/berumen_uncropped.png){#fig:fig_arms}

@@

```{r arms-table,results="asis", dependson="init"}
arms_schedule <- cc$crabs_untransformed %>%
  sample_tibble() %>%
  separate_wider_regex(sample,patterns=c(recovery_year="^[0-9]+","_",sample=".*"))  %>%
  select(sample,everything()) %>%
  unite("range",deployment_year,recovery_year,sep="--") %>%
  mutate(
    island = case_when(
      str_remove_all(hawaiian_name,'[^a-zA-Z ]') != island ~ as.character(str_glue("{hawaiian_name} ({island})")),
      TRUE ~ as.character(hawaiian_name)
    ),
    island = fct_reorder(island,-lat)
  ) %>%
  count(island,range) %>%
  pivot_wider(names_from="range",values_from="n")  %>%
  select(island,order(colnames(.))) %>%
  mutate(across(-island,~replace_na(.x,0)))
print_table(arms_schedule,"arms_table","ARMS unit sampling schedule (number of units)") 
```

```{r site-map, fig.cap='Sampling map (Papahānaumokuākea Marine National Monument shaded in light blue)', results="asis", dev="svg", fig.width=9, fig.height=6, dependson="init"}
library(sf)
library(rnaturalearth)

@@

ggplot() +
  # geom_contour(data=bath_proj,mapping=aes(x=V1,y=V2,z = V3),  color="#cecece") +
  # geom_raster(data=bpr,aes(x=x,y=y,fill=z)) +
  # geom_tile(data=bath_proj,mapping=aes(x=V1,y=V2,fill = V3)) +
  # scale_fill_gradient2(low="dodgerblue4", mid="gainsboro", high="darkgreen") +
  geom_sf(data=hawaii %>% st_transform(transformer), color="black", fill="green", linewidth=0.4) +
  geom_sf(data=pmnm,fill="#6ed8f6",alpha=0.1) +
  geom_sf(data=crab_sites %>% st_transform(transformer),aes(fill=hawaiian_name),size=3,shape=21,color="black") +
  scale_fill_manual(values=island_pal,name="Island") +
  coord_sf(xlim=box_after[c(1,3)],ylim=box_after[c(2,4)]) +
  scale_x_continuous(breaks = -seq(180, 155, -5)) +
  theme_minimal() +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(legend.text = element_text(size=10),legend.title = element_text(size=12))
cat("{#fig:site_map}")
```

ARMS units were retrieved by encapsulating them in-situ in a mesh-lined crate (100 μm pore size) to prevent escape of motile organisms. Once returned to the surface, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate. Seawater surrounding disassembled units was sieved through 2 mm, 500 μm, and 100 μm mesh-size geologic sieves [see @lerayDNABarcodingMetabarcoding2015 for processing details], and retained macroorganisms were preserved in 95% ethanol. All adult brachyuran crabs ($\geq$ 5 mm) were photographed, identified to morphospecies, and preserved in 95% ethanol (Fig. 2d). Individual preserved specimens were accessioned into the invertebrate zoology collections at the Florida Museum of Natural History (catalog numbers xxx–xxx).

## Data analysis {-}

@@

## Environmental variables {-}

Potential environmental factors affecting the community structure of brachyuran crabs on shallow reefs were assembled from published literature [@selkoeDNACoralReef2016], NOAA's National Coral Reef Monitoring Program (NCRMP), and the Pacific Islands Ocean Observing System (PacIOOS) ([@tbl:env_table]). Environmental factors were assembled only for shallow sites sampled during RAMP cruises.

```{r env-table,results="asis", dependson="init"}
print_table(here("data","tables","environmental_variables.csv"),"env_table","Potential environmental drivers of crab community structure, assembled from PacIOOS and @selkoeDNACoralReef2016")
```

# Results {-}

## Diversity summary {-}

### All ARMS {-}

The `r nsamples(cc$crabs_untransformed)` ARMS units deployed across the Hawaiian Islands (in `r format_range(cc$crab_data$depth)` m depth) yielded a total of `r num(sum(sample_sums(cc$crabs_untransformed)),1)` adult crabs. Sampling units contained `r num(all_summary$abundance_mean,0.1)` $\pm$ `r num(all_summary$abundance_sd,0.1)` (mean $\pm$ SD) individuals per unit, and abundance ranged from `r format_range(all_summary$abundance_range)` individuals. Crabs were identified to `r sum(unlist(all_summary$types_breakdown))` distinct types (including `r all_summary$types_breakdown$species` species and `r all_summary$types_breakdown$morphospecies` morphospecies) in `r all_summary$taxonomy_breakdown$genera` genera and `r all_summary$taxonomy_breakdown$families` families, which accounted for ~36% of the known crab diversity reported from Hawaiian waters [@castroCatalogAnomuranBrachyuran2011]. The four most common species were: `r with(all_summary,top_list(top5,top5_n,top5_units,fig="fig_crabs"))`. `r str_to_sentence(as.english(all_summary$singleton_count))` species occurred in a single unit across all sites (singletons), accounting for ~`r num(all_summary$singleton_count/ntaxa(cc$crabs_untransformed)*100,1)`% of total recovered crab species.

![The four most abundant crab species found across all deployed ARMS units. (A) *Chlorodiella laevissima*. (B) *Perinia tumida*. (C) *Percnon abbreviatum*. (D) *Dynomene hispida*.](data/images/4crabs.png){#fig:fig_crabs}

### Mesophotic ARMS {-}

@@

## Alpha diversity {-}

### Summary {-}

@@

### Comparisons {-}

For the complete dataset, there was no significant difference in observed species richness between the MHI and NWHI `r t_test(all_alpha$t_r_island)`, but Simpson diversity was significantly different, although mean values were similar with `r num(all_alpha$richness_main$richness,0.1)` $\pm$ `r num(all_alpha$richness_main$richness_sd,0.1)` in the MHI and `r num(all_alpha$richness_nwhi$richness,0.1)` $\pm$ `r num(all_alpha$richness_nwhi$richness_sd,0.1)` in the NWHI ([@fig:alpha_box]; `r t_test(all_alpha$t_s_island,plain=TRUE)`). When only shallow arms \< 30 m were considered, neither comparison was significant (richness: `r t_test(shallow_alpha$t_r_island,plain=TRUE)`; Simpson: `r t_test(shallow_alpha$t_s_island,plain=TRUE)`).

```{r alpha-box, fig.cap='Box plots of Simpson diversity by island group. Notches show nonparametric confidence intervals about the median', results="asis", dev="svg", fig.width=10, fig.height=8, dependson="init"}
ggplot(all_alpha$richness_table %>% mutate(island_group = str_to_sentence(island_group))) + 
  geom_boxplot(aes(x=island_group,y=Simpson,fill=island_group),notch=TRUE) +
  scale_fill_manual(values=island_group_pal) +
  guides(fill="none") + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("Island group") + 
  ylab("Simpson's diversity index") 
cat("{#fig:alpha_box}")
```

### Effects of environmental variables {-}

## Brachyuran community structure {-}

```{r permanova, dependson="init"}
# PERMANOVA of 
anova_region <- adonis2(cc$crab_dist_shallow ~ island_group,data=cc$crab_data_shallow,permutations=9999)
anova_island <- adonis2(cc$crab_dist_shallow ~ island,data=cc$crab_data_shallow,permutations=9999)
```

There were significant differences in shallow crab community structure between the Main (MHI) and Northwest Hawaiian Islands (`adonis2` PERMANOVA: `r permanova(anova_region)`) as well as among individual islands (`r permanova(anova_island)`; see [@fig:pairwise_island] for results of pairwise comparisons).

# Discussion {-}

# References {-}

@@

# Supplemental material {label="S"}

```{r pairwise-island, fig.cap='Results of pairwise PERMANOVA tests by island (FDR-adjusted p-values)', results="asis", dev="svg", fig.width=12, fig.height=9, dependson="init"}
library(ggtext)
pairwise_island <- pairwise_adonis(cc$crab_dist_shallow,cc$crab_data_shallow$island,permutations=999)

pw <- pairwise_island %>%
  filter(p_adj < 0.05)

islands <- cc$crabs_shallow %>%
  sample_tibble() %>%
  pull(island) %>%
  levels()

pw1 <- pairwise_island %>%
  select(left,right,p_adj) %>%
  mutate(left=factor(left,levels=rev(islands)),right=factor(right,levels=rev(islands))) %>%
  mutate(sig = p_adj < 0.05, p_adj = pval(p_adj)) %>%
  mutate(
    combo = map2_chr(left,right,~str_c(sort(c(.x,.y)),collapse="-"))
  ) %>%
  select(-c(left,right)) %>%
  separate_wider_delim(combo,"-",names=c("left","right")) %>%
  select(left,right,everything()) %>%
  mutate(left=factor(left,levels=islands),right=factor(right,levels=islands))

ggplot(pw1) + 
  # geom_tile(aes(x=right,y=left,fill=sig),width=0.6,height=0.8) + 
  # geom_tile(aes(x=right,y=left),width=0.6,height=0.8) + 
  geom_richtext(
    aes(
      x=right,
      y=left,
      label=p_adj,
      fontface = if_else(sig,"bold","plain")
    ),
    fill = NA,
    label.color = NA
  ) +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color="black")
  ) +
  # scale_fill_manual(values = c("white","lightgoldenrod")) +
  guides(fill="none") +
  xlab("") + ylab("") +
  scale_x_discrete(position="top") +
  scale_y_discrete(limits=rev)
cat("{#fig:pairwise_island}")
```
